---
title: "EDA Analysis Beer Store"
author: "Carlos Estevez"
date: "2023-02-17"
output: 
      html_document:
        number_sections: true
---

```{r Libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(caret)
library(plotly)
library(ggthemes)
library(GGally)
library(class)
library(e1071)
library(stringr)
```

# Introduction

This analysis is about the number of beers consumed by American People in the United States
In this docoument we conduct an analysis about 


### Reading data files

```{r ReadingFiles, warning=FALSE,message=FALSE,include=FALSE}

setwd("C:/Users/cestevez/Dropbox/Cloud PC/Thinkpad/Thinkpad Desktop/Master Data Science SMU/Class_Sessions/Data Science Sessions/Repository/SMU_MSDS_6306")

df_beers = read.csv("session_8/Beers.csv",header = TRUE)
df_breweries_1 = read.csv("session_8/Breweries.csv",header = TRUE)
df_states = read.csv("session_8/States.csv",header=TRUE)
df_breweries_1$State = str_trim(df_breweries_1$State)
df_breweries_2 = merge(df_breweries_1,df_states,by = "State")

```

### Addressing the missing values in each column

```{r Filtering_Data}

#Cleaning ABV using mean
df_beers_cl0 = df_beers
nr_mean_abv = mean(df_beers_cl0[!is.na(df_beers_cl0$ABV),]$ABV)
length_abv = length(df_beers_cl0[is.na(df_beers_cl0$ABV),]$ABV)
if(length_abv > 0){
  df_beers_cl0[is.na(df_beers_cl0$ABV),]$ABV = nr_mean_abv
}

#Cleaning IBU using mean
df_beers_cl1 = df_beers
nr_mean_ibu = mean(df_beers_cl1[!is.na(df_beers_cl1$IBU),]$IBU)
length_ibu = length(df_beers_cl1[is.na(df_beers_cl1$IBU),]$IBU)
if(length_ibu > 0){
  df_beers_cl1[is.na(df_beers_cl1$IBU),]$IBU = nr_mean_ibu
}


#Cleaning using KnnInpute
# preProcValues <- preProcess(df_beers %>% 
#                           select(ABV,IBU),
#                             method = c("knnImpute"),
#                             k = 20,
#                             knnSummary = mean)
# df_beers_unp <- predict(preProcValues, df_beers,na.action = na.pass)
# procNames <- data.frame(col = names(preProcValues$mean), mean = preProcValues$mean, sd = preProcValues$std)
# for(i in procNames$col){
#  df_beer_info[i] <- df_beer_info[i]*preProcValues$std[i]+preProcValues$mean[i] 
# }
knn_imp_model <- preProcess(df_beers_cl0 %>%
                          select(ABV,IBU),
                            method = c("knnImpute"),
                            k = 20,
                            knnSummary = mean)



df_beers_unp <- predict(knn_imp_model, df_beers_cl0,na.action = na.pass)
procNames <- data.frame(col = names(knn_imp_model$mean), mean = knn_imp_model$mean, sd = knn_imp_model$std)
for(i in procNames$col){
 df_beers_unp[i] <- df_beers_unp[i]*knn_imp_model$std[i]+knn_imp_model$mean[i] 
}


#Cleansing beers, records 2410
nr_rows = dim(df_beers)[1]
#Finding the NAN values
#summary(df_beers)
# df_beers  %>% ggplot(aes(x=IBU))+geom_histogram(aes(fill="green")) 
# df_beers  %>% ggplot(aes(x=ABV))+geom_histogram(aes(fill="green")) 
# 
# df_beers_cl1  %>% ggplot(aes(x=IBU))+geom_histogram(aes(fill="blue"))
# df_beers_cl1  %>% ggplot(aes(x=ABV))+geom_histogram(aes(fill="blue"))

df_beers_unp  %>% ggplot(aes(x=IBU))+geom_histogram() 
df_beers_unp  %>% ggplot(aes(x=ABV))+geom_histogram() 
df_beers_unp %>% ggplot(aes(y=ABV))+geom_boxplot()

summary(df_beers_unp)


```

### How many breweries are present in each state

```{r Summarizing by State}
df_summary = df_breweries_2 %>% group_by(State,Name_State) %>% summarize(NumberBreweries = n())

knitr::kable(
  df_summary,
  caption = "Number of Beers by State"
)

```

### Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file

```{r Merging_Data}
df_beerbre_unp = merge(df_beers_unp,df_breweries_2,by.x = "Brewery_id",by.y = "Brew_ID")

```

### Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare

```{r Median_Beer_Brewer, message=FALSE}

df_acom_bebrew_1 = df_beerbre_unp %>% group_by(State,Name_State) %>% summarize(Median_ABV = mean(ABV),Median_IBU = mean(IBU))

# df_acom_bebrew_1 %>% ggplot(aes(x=State,color=State))+geom_bar()+labs(title = "Alcohol by Volume",subtitle = "Alcohol by Volume average by State")+coord_flip()

```
### Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer

```{r}

df_sort_1 = arrange(df_beerbre_unp,desc(ABV)) %>% head(n = 1)
sprintf("The state that has the maximum ABV is %s-%f",df_sort_1$Name_State,df_sort_1$ABV)

```
### Comment on the summary statistics and distribution of the ABV variable
#### Accordint to the histogram we can notice that the data seems normally distributed
```{r}

df_beerbre_unp %>% ggplot(aes(x=ABV,fill=State))+geom_histogram()

```
### Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer

```{r}
df_beerbre_unp %>% select(IBU,ABV) %>% ggpairs(columnLabels = c("ABV","IBU"))



```


###Activity 8, Difference with respect to IBU and ABV IPA and ALE

```{r}

df_beerbre_fil0 = filter(df_beerbre_unp,str_detect(df_beerbre_unp$Style,regex("IPA|ALE",ignore_case = TRUE)))

df_beerbre_fil1 = mutate(df_beerbre_fil0,Type= ifelse(str_detect(Style,regex("IPA",ignore_case = TRUE)),"IPA","ALE"))


df_beerbre_fil1 %>% ggplot(aes(x=ABV,y=IBU,color=Type))+geom_point()+
  labs(title = "Relationship IBU and AVB",subtitle = "Relationship IBU/AVB by Beer Type")+xlab("Alcohol by Volume (ABV)")+ylab("International Bitterness Unit(IBU)")


# df_beerbre_fil1 = filter(df_beerbre_unp,str_detect(df_beerbre_unp$Style,regex("IPA",ignore_case = TRUE)))
# 
# df_beerbre_fil2 = filter(df_beerbre_unp,str_detect(df_beerbre_unp$Style,regex("ALE",ignore_case = TRUE)))

nr_percentage = 0.7
nr_observations = nrow(df_beerbre_fil1)
nr_k = 5

lst_index_beer = sample(nr_observations,round(nr_observations * nr_percentage))
df_train = df_beerbre_fil1[lst_index_beer,]
df_test = df_beerbre_fil1[-lst_index_beer,]

knn_result = knn(df_train[,c(4,5)],df_test[,c(4,5)],df_train$Type,prob = TRUE,k=nr_k)
co_table =  table(knn_result,df_test$Type)
confusionMatrix(co_table)

#When it comes down to it, a Pale Ale should have a nice hop character but medium build, whereas IPAs tend to have higher ABV and IBU.







```
```{r}

df_sum_1 = group_by(df_beerbre_unp,State,Name_State,Style) %>% summarize(count_style = n())
df_sum_1 = mutate(df_sum_1,flavor_type = ifelse(str_detect(Style,regex("IPA",ignore_case = TRUE)),"Ipa","Pale"))
df_sum_1 = mutate(df_sum_1,general_stype = ifelse(str_detect(Style,regex("lager",ignore_case = TRUE)),"Lager","Ale"))

df_sum_1 %>% ggplot(aes(x=Name_State,y=count_style,fill=flavor_type))+geom_bar(stat = "identity",position = "stack")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



df_sum_2 = group_by(df_beerbre_unp,State,Name_State,Style) %>% summarize(count_style = n()) %>% filter(count_style>5) %>% mutate(Name_State_2=str_trim(str_to_lower(Name_State))) 

df_sum_4 = group_by(df_beerbre_unp,State,Name_State,Style) %>% summarize(count_style = n()) %>%  mutate(Name_State_2=str_trim(str_to_lower(Name_State)),Lager=ifelse(str_detect(Style,regex("lager",ignore_case = TRUE)),"X","")) %>%
  filter(count_style>5)



df_sum_3 = group_by(df_beerbre_unp,State,Name_State) %>% summarize(count_style = n()) %>% mutate(Name_State_2=str_trim(str_to_lower(Name_State))) 


options(repr.plot.width = 1000, repr.plot.height =2)

gfr = df_sum_2 %>% arrange(Name_State) %>% ggplot(aes(x=Style,y=count_style,fill = Name_State ))+geom_bar(stat = "identity",position = "stack")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),legend.position = "left",text = element_text(size = 10))

ggplotly(gfr)


 df_sum_4 %>% arrange(Name_State) %>% ggplot(aes(x=Name_State,y=count_style,fill = Style))+geom_bar(stat = "identity",position = "stack")+geom_text(aes(label = Lager), vjust = 1.5, colour = "white")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),legend.position = "left",text = element_text(size = 10))

#ggplotly(gfr)



library(maps)
usa = map_data("state")
all_states <- map_data("state")


all_states_2 = all_states %>% group_by(region) %>% summarize(lat=mean(lat),long=mean(long)) %>% mutate(AB_State = str_to_upper(str_sub(region,1,2)))


state_com_data = merge(df_sum_2,all_states_2,by.x = "Name_State_2",by.y = "region")



# all_states %>% ggplot()+geom_polygon(aes(x=long, y=lat, group = group,fill=region))+theme(legend.position = "none")+geom_text(data=all_states_2,aes(x=long, y=lat,label=region))


all_states_3 = merge(all_states,df_sum_3,by.x="region",by.y = "Name_State_2")

all_states_4 = merge(all_states_2,df_sum_3,by.x="region",by.y = "Name_State_2") 



# all_states_3 %>% ggplot()+geom_polygon(aes(x=long, y=lat, group = group,fill=count_style))+theme(legend.position = "none")+geom_label(data=all_states_2,aes(x=long, y=lat,label=AB_State,alpha=0.5))

# all_states %>% ggplot()+geom_polygon(aes(x=long, y=lat, group = group,fill=region))+theme(legend.position = "none")+geom_label(data=all_states_2,aes(x=long, y=lat,label=sprintf("asdsadsa\ndasd"),alpha=0.5))

all_states %>% ggplot()+geom_polygon(aes(x=long, y=lat, group = group,fill=region))+theme(legend.position = "none")+
  ggrepel::geom_label_repel(aes(x=long, y=lat,label=sprintf("%s,Beers:%i",Name_State,count_style),alpha=0.8),size=2,data=all_states_4)

```

