---
title: "Practicing Regex"
author: "Carlos Estevez"
date: "2023-02-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Practicing with String Functions

```{r PracticingFunction }
library(stringr)

str1 = "This is the way it will \"print out\""
vect_1 = c("Hello","This is an example")
str2 = "The number of beers"
str3 = "300 units"

#WriteLines
print(str1)
writeLines(str1)

#StrLength 
nr_leng = str_length(vect_1)
nr_leng 

#Combining Strings
result_1 = str_c(str2,str3,sep = " ")
result_1

#Substring

st_beers = str_sub(result_1,21,23)
st_beers

#Upper case and lower case

str_to_lower(str1)
str_to_upper(str1)

#Trim and Wrap
str_trim(str1) 

str_4 = "The str_wrap() function gives us a way to wrap strings into paragraphs. The formatting of the paragraphs can be adjusted by width and can include an indent or exdent. str_wrap() is typically combined with the cat() (concatenate) function"

str_5 = str_wrap(str_4,width = 20)
writeLines(str_5)



```
## Practicing with Regex
```{r}

#Using str_view and Str_View_all
x1 = c("banana","apple","pear")
x2 = c("action","reaction","This is the Pollution","No way action will be found","No one")
str_view(x1,"an") #It will find all patterns related to an
str_view_all(x1,"an")
str_view_all(x2,"ion")
str_view_all(x2,"ion.") #This is different from the last statement

#Using Anchors

str_view_all(x2,"^a")
str_view_all(x2,"e$")
str_view_all(x2,"^action$")
str_view_all(x2,"^Pollution$")

#We want to filter only the word action
str_view_all(x2,"action")  #The system will bring action and reaction
str_view_all(x2,"\\baction\\b")


#Matching any digit
str_6 = c("This is the number 1","No a number")
str_view_all(str_6,"\\d")
str_view_all(str_6,"\\s")




```
## Or or Alternative Path
```{r}
str_7 = c("action abc","This is the way xyz","Nothing is special")
str_8 = c("charles","cherles","Jose")

str_view_all(str_7,"abc|xyz")
str_view_all(str_8,"ch(a|e)rles")


```
## Repetition
* A character class is a list of characters enclosed between ‘⁠[⁠’ and ‘⁠]⁠’ which matches any single character in that list; unless the first character of the list is the caret ‘⁠^⁠’, when it matches any character not in the list. For example, the regular expression ‘⁠[0123456789]⁠’ matches any single digit, and ‘⁠[^abc]⁠’ matches anything except the characters ‘⁠a⁠’, ‘⁠b⁠’ or ‘⁠c⁠’
* Alphanumeric characters: ‘⁠[:alpha:]⁠’ and ‘⁠[:digit:]⁠’.
* ⁠[:blank:]⁠’
Blank characters: space and tab, and possibly other locale-dependent characters, but on most platforms not including non-breaking space.
* ⁠[:alpha:]⁠’
Alphabetic characters: ‘⁠[:lower:]⁠’ and ‘⁠[:upper:]
⁠
* [:alnum:]
Alphanumeric characters: ‘⁠[:alpha:]⁠’ and ‘⁠[:digit:]
⁠[:lower:]
Lower-case letters in the current locale.
* ‘⁠[:digit:]⁠’
Digits: ‘⁠0 1 2 3 4 5 6 7 8 9⁠’.

```{r}
str_9 = c("CCdsadCC","CC","CCCCAA")
str_view_all(str_9,"C?") #?-->0 or 1, +: 1 or more, *: 0 or more
str_view_all(str_9,"CC{1}")

#How to validate a phone number dominican republic
str_9 = "\\d{3}-\\d{3}-\\d{4}"
str_10 = c("809-222-4969","5541")
str_view_all(str_10,str_9)

#Find all the words that start with three consonants
str_11 = c("string","carlos","stronger")
str_12 = "[^aeiou]{3}"
str_view_all(str_11,str_12)


#Here we want a word that end with ed and not eed
str13 = "([^e])ed$"
str14 = c("checaeed","eddcheck","tired")
str_view_all(str14,str13)

#is Q always followed by a U
str_15 = "qu"

#Word that has two voews or more

str_16 = c("actioon","repetition")
str_17 = "[aeiou]{3,}"
str_view_all(str_16,str_17)

```
## Detect Matches

```{r}

str_16 = "tion|ed$"
lst_1 = c("Reaction","Action","Tired","Bored","Carlos","Leverage")
str_detect(lst_1,str_16)
lst_1[str_detect(lst_1,str_16)]



```

# Extracting data
* What the function str_extract will do is to extract the matched part of the word

[abc]: matches a, b, or c.
[a-z]: matches every character between a and z (in Unicode code point order).
[^abc]: matches anything except a, b, or c.
[\^\-]: matches ^ or -.
[:punct:]: punctuation.
[:alpha:]: letters.
[:lower:]: lowercase letters.
[:upper:]: upperclass letters.
[:digit:]: digits.
[:xdigit:]: hex digits.
[:alnum:]: letters and numbers.
[:cntrl:]: control characters.
[:graph:]: letters, numbers, and punctuation.
[:print:]: letters, numbers, punctuation, and whitespace.
[:space:]: space characters (basically equivalent to \s).
[:blank:]: space and tab.
```{r}
str_16 = "tion|ed$"
lst_1 = c("Reaction","Action","Tired","Bored","Carlos","Leverage")
#str_extract(lst_1,str_16)
str_match(lst_1,str_16)
#str_extract_all(lst_1,str_16)



```

# Repetition and Counting
```{r}

lst_t1 = c("809-898-9990","2344-7799-8669","444-123-1112")
str_17 = "\\d{3}-\\d{3}-\\d{4}"
str_detect(lst_t1,str_17)

str_count(lst_t1,str_17)

```
# Replaning characters
* The system will replace the entire string for the replacing character
```{r}

lst_t1 = c("809-898-9990","2344-7799-8669","444-123-1112")
str_17 = "\\d{3}-\\d{3}-\\d{4}"
str_replace_all(lst_t1,str_17,"444")
str_replace(lst_t1,str_17,"444")


str_18 = "(a) (c)"
str_detect("a crlos",str_18) #This is TRUE

str_18 = "([^a]) (c)"
str_detect("b crlos",str_18) #This is TRUE

str_19 = "([^a]) (c)"
str_detect("b crlos",str_18) #This is TRUE


str_20 = "(a+) (c)"
str_detect("aba crlos",str_20) #This is TRUE, what we are saying here is A has to repeat multiple times

str_21 = "c(a)rlos"
str_22 = "carlos"
lst_t2 = c("carlos")
str_detect(lst_t2,str_21) #Both are TRUE




```
# Splitting

```{r}

str = "This is a great adventure"
unlist(str_split(str," "))


str = "This,is,a,great,adventure"
str_split(str,",",simplify = TRUE)

str = "This,is,a,great,adventure"
str_split(str,boundary("word"),simplify = TRUE)




```
# Find matches
```{r}
str = "This,is,a,great,adventure This"
str_locate(str,"This")
str_locate_all(str,regex("this",ignore_case = TRUE))




```

#Subset 

```{r}

str = c("This,is,a,great,adventure This","carlos")
str_subset(str,"This")
str_extract(str,"This")



```
# Stringi
```{r}



```









