---
title: "EDA Analysis Beer Store"
author: "Carlos Estevez"
date: "2023-02-17"
output: 
      html_document:
        number_sections: true
---

```{r Libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(caret)
library(plotly)
library(ggthemes)
library(GGally)
library(class)
library(e1071)
```

# Introduction

This analysis is about the number of beers consumed by American People in the United States

###Reading data files

```{r ReadingFiles, warning=FALSE,message=FALSE,include=FALSE}

setwd("C:/Users/cestevez/Dropbox/Cloud PC/Thinkpad/Thinkpad Desktop/Master Data Science SMU/Class_Sessions/Data Science Sessions/Repository/SMU_MSDS_6306")

df_beers = read.csv("session_8/Beers.csv",header = TRUE)
df_breweries_1 = read.csv("session_8/Breweries.csv",header = TRUE)
df_states = read.csv("session_8/States.csv",header=TRUE)
df_breweries_1$State = str_trim(df_breweries_1$State)
df_breweries_2 = merge(df_breweries_1,df_states,by = "State")

```

### Addressing the missing values in each column

```{r Filtering_Data}

#Cleaning ABV using mean
df_beers_cl1 = df_beers
nr_mean_abv = mean(df_beers_cl1[!is.na(df_beers_cl1$ABV),]$ABV)
length_abv = length(df_beers_cl1[is.na(df_beers_cl1$ABV),]$ABV)
if(length_abv > 0){
  df_beers_cl1[is.na(df_beers_cl1$ABV),]$ABV = nr_mean_abv
}

#Cleaning IBU using mean
nr_mean_ibu = mean(df_beers_cl1[!is.na(df_beers_cl1$IBU),]$IBU)
length_ibu = length(df_beers_cl1[is.na(df_beers_cl1$IBU),]$IBU)
if(length_ibu > 0){
  df_beers_cl1[is.na(df_beers_cl1$IBU),]$IBU = nr_mean_ibu
}


#Cleaning using KnnInpute
# preProcValues <- preProcess(df_beers %>% 
#                           select(ABV,IBU),
#                             method = c("knnImpute"),
#                             k = 20,
#                             knnSummary = mean)
# df_beers_unp <- predict(preProcValues, df_beers,na.action = na.pass)
# procNames <- data.frame(col = names(preProcValues$mean), mean = preProcValues$mean, sd = preProcValues$std)
# for(i in procNames$col){
#  df_beer_info[i] <- df_beer_info[i]*preProcValues$std[i]+preProcValues$mean[i] 
# }
knn_imp_model <- preProcess(df_beers %>%
                          select(ABV,IBU),
                            method = c("knnImpute"),
                            k = 20,
                            knnSummary = mean)



df_beers_unp <- predict(knn_imp_model, df_beers,na.action = na.pass)
procNames <- data.frame(col = names(knn_imp_model$mean), mean = knn_imp_model$mean, sd = knn_imp_model$std)
for(i in procNames$col){
 df_beers_unp[i] <- df_beers_unp[i]*knn_imp_model$std[i]+knn_imp_model$mean[i] 
}


#Cleansing beers, records 2410
nr_rows = dim(df_beers)[1]
#Finding the NAN values
summary(df_beers)
df_beers  %>% ggplot(aes(x=IBU))+geom_histogram(aes(fill="green")) 
df_beers  %>% ggplot(aes(x=ABV))+geom_histogram(aes(fill="green")) 
df_beers_cl1  %>% ggplot(aes(x=IBU))+geom_histogram(aes(fill="blue"))
df_beers_cl1  %>% ggplot(aes(x=ABV))+geom_histogram(aes(fill="blue"))
df_beers_unp  %>% ggplot(aes(x=IBU))+geom_histogram() 
df_beers_unp  %>% ggplot(aes(x=ABV))+geom_histogram() 

df_beerbre_unp %>% ggplot(aes(y=ABV))+geom_boxplot()




```

### How many breweries are present in each state

```{r Summarizing by State}
df_summary = df_breweries_2 %>% group_by(State,Name_State) %>% summarize(NumberBreweries = n())

knitr::kable(
  df_summary,
  caption = "Number of Beers by State"
)

```

### Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file

```{r Merging_Data}
df_beerbre_unp = merge(df_beers_unp,df_breweries_2,by.x = "Brewery_id",by.y = "Brew_ID")

```

### Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare

```{r Median_Beer_Brewer, message=FALSE}

df_acom_bebrew_1 = df_beerbre_unp %>% group_by(State,Name_State) %>% summarize(Median_ABV = mean(ABV),Median_IBU = mean(IBU))

# df_acom_bebrew_1 %>% ggplot(aes(x=State,color=State))+geom_bar()+labs(title = "Alcohol by Volume",subtitle = "Alcohol by Volume average by State")+coord_flip()

```
### Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer

```{r}

df_sort_1 = arrange(df_beerbre_unp,desc(ABV)) %>% head(n = 1)
sprintf("The state that has the maximum ABV is %s-%f",df_sort_1$Name_State,df_sort_1$ABV)

```
### Comment on the summary statistics and distribution of the ABV variable
#### Accordint to the histogram we can notice that the data seems normally distributed
```{r}

df_beerbre_unp %>% ggplot(aes(x=ABV,fill=State))+geom_histogram()

```
### Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer

```{r}
df_beerbre_unp %>% select(IBU,ABV) %>% ggpairs(columnLabels = c("ABV","IBU"))



```


###Activity 8, Difference with respect to IBU and ABV IPA and ALE

```{r}

df_beerbre_fil0 = filter(df_beerbre_unp,str_detect(df_beerbre_unp$Style,regex("IPA|ALE",ignore_case = TRUE)))

df_beerbre_fil1 = mutate(df_beerbre_fil0,Type= ifelse(str_detect(Style,regex("IPA",ignore_case = TRUE)),"IPA","ALE"))


df_beerbre_fil1 %>% ggplot(aes(x=ABV,y=IBU,color=Type))+geom_point()+
  labs(title = "Relationship IBU and AVB",subtitle = "Relationship IBU/AVB by Beer Type")+xlab("Alcohol by Volume (ABV)")+ylab("International Bitterness Unit(IBU)")


# df_beerbre_fil1 = filter(df_beerbre_unp,str_detect(df_beerbre_unp$Style,regex("IPA",ignore_case = TRUE)))
# 
# df_beerbre_fil2 = filter(df_beerbre_unp,str_detect(df_beerbre_unp$Style,regex("ALE",ignore_case = TRUE)))

nr_percentage = 0.7
nr_observations = nrow(df_beerbre_fil1)
nr_k = 5

lst_index_beer = sample(nr_observations,round(nr_observations * nr_percentage))
df_train = df_beerbre_fil1[lst_index_beer,]
df_test = df_beerbre_fil1[-lst_index_beer,]

knn_result = knn(df_train[,c(4,5)],df_test[,c(4,5)],df_train$Type,prob = TRUE,k=nr_k)
co_table =  table(knn_result,df_test$Type)
confusionMatrix(co_table)

#When it comes down to it, a Pale Ale should have a nice hop character but medium build, whereas IPAs tend to have higher ABV and IBU.







```

